---
title: "MakeTheFounderPopulation"
author: 
  name: "Jean-Luc Jannink"
  affiliation: "USDA-ARS / Cornell"
date: "September 7th, 2024"
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    highlight: haddock
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
      toc_depth: 2
editor_options: 
  chunk_output_type: console
---
  
<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

Create the founder population that will be the basis for within-family
prediction The assumption is that different founder populations might behave
somewhat differently

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(cache=FALSE)
```

#### Load packages first  
```{r load packages, message=TRUE}
packages_used <- c("AlphaSimR", "tidyverse", "workflowr", "here", "rrBLUP",
                   "EMMREML")
ip <- installed.packages()
all_packages_installed <- TRUE
for (package in packages_used){
  if (!(package %in% ip[,"Package"])){
    print(paste("Please install package", package))
    all_packages_installed <- FALSE
  }
}#END packages_used
if (!all_packages_installed) stop("Need to install packages")

library(tidyverse)
```

#### Set file locations relative to the project folder using `here`  
```{r set here path}
here::i_am("analysis/MakeTheFounderPopulation.Rmd")
```

#### Document packages used (sessionInfo)  
```{r print Session Info}
print(utils::sessionInfo())
```

#### Set random seed  
```{r set random seed}
random_seed <- 45678
set.seed(random_seed)
```

#### Script parameters  
If the behavior of your script depends on parameters that you set, 
initialize them early on.  
```{r simulation parameters}
nChr <- 1 # Number of chromosomes
nSnpPerChr <- 200 # Number of segregating sites _per chromosome_
nQTLperChr <- 20 # Vary this parameter to get oligo- versus poly- genic traits

nFounders <- 1000
nProgeny <- 100
```

#### AlphaSimR population  
```{r Make founder population}
# Create haplotypes for founder population of outbred individuals
# Note: default effective population size for runMacs is 100
founderHaps <- AlphaSimR::runMacs(species="WHEAT", nInd=nFounders, nChr=nChr, 
                                  segSites=nSnpPerChr+nQTLperChr)

# New global simulation parameters from founder haplotypes
SP <- AlphaSimR::SimParam$new(founderHaps)
SP$restrSegSites(nQTLperChr, nSnpPerChr, overlap=FALSE)
SP$addSnpChip(nSnpPerChr)

# Additive trait architecture
# By default, the genetic variance will be 1
SP$addTraitA(nQtlPerChr=nQTLperChr)

# Create a new population of founders
founders <- AlphaSimR::newPop(founderHaps, simParam=SP)
# Make the founders inbred
founders <- AlphaSimR::makeDH(founders)
```

```{r Get to here}

```

