---
title: "MakeTheFounderPopulation"
author: 
  name: "Jean-Luc Jannink"
  affiliation: "USDA-ARS / Cornell"
date: "September 7th, 2024"
output:
  html_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 12
    highlight: haddock
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: no
      toc_depth: 2
editor_options: 
  chunk_output_type: console
---
  
<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>

Create the founder population that will be the basis for within-family
prediction The assumption is that different founder populations might behave
somewhat differently

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
knitr::opts_chunk$set(message=FALSE)
knitr::opts_chunk$set(cache=FALSE)
```

## Taking the marker perspective to optimize the training population.  
Hypothesis: if an individual has a long stretch of IBS with a parent then it is
in fact IBD with the parent and it will better predict the parent effect in that
stretch.
1. Inbred individuals  
2. Each ind has a length IBS with each parent haplotype  
3. Maximize the sum of lengths  
4. Minimize the KS deviation from uniformity, done separately for each parent  
5. Minimize the difference in lengths between the two parents  

The basic idea is to construct a training population with these characteristics
by combinatorial optimization.  So pick a population and then iteratively 
improve it by swapping individuals and testing if a utility function increases.  

Ultimately, the objective is to construct a training population in which the 
LD relationships between loci are as similar as possible to what they would be
among the progeny of the two parents. Fulfilling points 3 to 5 aims to achieve
that while not requiring that you calculate those LD relationships at each 
iteration of the optimization.

Noteworthy
1. The optimization can lead to TPs that have notably worse prediction
accuracy than a random population. I have yet to construct TPs that have notably
better accuracy. But just the fact of variation across the parameters of the
optimization is hopeful.
2. There is quite a bit of variation in the accuracy achieved among repeated
optimizations. I have not yet looked at whether characterization of the
constructed TP can illuminate what causes the differences in the ultimate
accuracy.

Tasks
1. See if very preliminary results are reproduced when I have higher SNP and QTL
densities. Hmmm: I can see that that is substantially slower.  Also, not any
better.  The optimization parameters might need to be revised.
2. Start setting up the approach for multiple chromosomes.
3. Start setting up the approach where the optimization is actually on the LD
relationships among the markers.

#### Load packages first  
```{r load packages, message=TRUE}
packages_used <- c("AlphaSimR", "tidyverse", "workflowr", "here", "rrBLUP",
                   "EMMREML")
ip <- installed.packages()
all_packages_installed <- TRUE
for (package in packages_used){
  if (!(package %in% ip[,"Package"])){
    print(paste("Please install package", package))
    all_packages_installed <- FALSE
  }
}#END packages_used
if (!all_packages_installed) stop("Need to install packages")

library(tidyverse)
```

#### Set file locations relative to the project folder using `here`  
```{r set here path}
here::i_am("analysis/MakeTheFounderPopulation.Rmd")
```

#### Document packages used (sessionInfo)  
```{r print Session Info}
print(utils::sessionInfo())
```

#### Set random seed  
```{r set random seed}
random_seed <- 45678
set.seed(random_seed)
```

#### Script parameters  
If the behavior of your script depends on parameters that you set, 
initialize them early on.  
```{r simulation parameters}
nChr <- 1 # Number of chromosomes
nSnpPerChr <- 200 # Number of segregating sites _per chromosome_
nQTLperChr <- 20 # Vary this parameter to get oligo- versus poly- genic traits

nFounders <- 1000
nProgeny <- 100
```

#### AlphaSimR population  
```{r Make founder population}
# Create haplotypes for founder population of outbred individuals
# Note: default effective population size for runMacs is 100
founderHaps <- AlphaSimR::runMacs(species="WHEAT", nInd=nFounders, nChr=nChr, 
                                  segSites=nSnpPerChr+nQTLperChr)

# New global simulation parameters from founder haplotypes
SP <- AlphaSimR::SimParam$new(founderHaps)
SP$restrSegSites(nQTLperChr, nSnpPerChr, overlap=FALSE)
SP$addSnpChip(nSnpPerChr)

# Additive trait architecture
# By default, the genetic variance will be 1
SP$addTraitA(nQtlPerChr=nQTLperChr)

# Create a new population of founders
founders <- AlphaSimR::newPop(founderHaps, simParam=SP)
# Make the founders inbred
founders <- AlphaSimR::makeDH(founders)
```

```{r Get to here}

```

