---
title: "Probability that QTL is IIS conditional on a sequence of marker IIS"
author: "Jean-Luc Jannink"
date: "2024-09-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### What to run from other markdowns.  
From MakeTheFounderPopulation.Rmd
`setup`
`load packages`
`set here path`
`set random seed`
`simulation parameters`
`Make founder population`

1. Make the population.  
2. Repeat.  
  i. Grab two individuals, two chromosomes per individual.  
  ii. On each chromosome, find the minimum probability streak.  
  iii. If there is a QTL in that streak, is it IIS between the individuals?  
  iv. Are the probabilities of the lengths of the streaks across the chromosomes correlated?  

```{r Set here base}
here::i_am("analysis/ProbQTLIISasFuncOfMrkIIS.Rmd")
```

I have a two-chromosome population, so I might as well use that: saveGlobalEnvObj21234.rds.  
```{r Get GlobalEnv}
  globalEnvObj <- readRDS(here::here("output", "saveGlobalEnvObj72345.rds"))
  for (n in names(globalEnvObj)) assign(n, globalEnvObj[[n]], envir=.GlobalEnv)
```



```{r Loop to estimate this probability}
nInd <- AlphaSimR::nInd(founders)
source(here::here("code", "assessIndMrkSimilarityToPar.R"))
# probSameSame are all the probabilities needed to calculate the streak probability
# for the founders population
probSameSame <- calcProbSameSame(founders)

allRes <- NULL
for (indPair in 1:1000){
  # Pick two individuals from the founders
  twoInd <- sample(nInd, 2)
  for (chr in 1:2){
    mrkHaplos <- AlphaSimR::pullSnpHaplo(founders[twoInd], chr=chr)
    nMrk <- ncol(mrkHaplos)
    haplo1 <- mrkHaplos[1,]; haplo2 <- mrkHaplos[3,]
    probWhere <- findPosMinProbSame(haplo1, haplo2, probSameSame[[chr]])
    # Figure out the actual IIS streak to determine if there is a QTL there
    mrkSame <- haplo1 == haplo2
    endPos <- probWhere[2]
    while (mrkSame[endPos+1]){
      endPos <- endPos+1
      if (endPos == nMrk) break
    }
    snpMap <- AlphaSimR::getSnpMap()
    strtPos <- snpMap$pos[probWhere[2]]
    endPos <- snpMap$pos[endPos]
    # Are there QTL on the streak?
    qtlMap <- AlphaSimR::getQtlMap()
    qtlMap <- qtlMap[qtlMap$chr == chr,]
    qtlInStreak <- which(qtlMap$pos >= strtPos & qtlMap$pos <= endPos)
    if (length(qtlInStreak) > 0){
      qtlRes <- rep(probWhere[1], length(qtlInStreak))
      qtlHaplos <- AlphaSimR::pullQtlHaplo(founders[twoInd], chr=chr)
      qtlHaplos <- qtlHaplos[c(1,3), qtlInStreak, drop=F]
      qtlRes <- cbind(qtlRes, qtlHaplos[1,] == qtlHaplos[2,])
      allRes <- rbind(allRes, qtlRes)
    }
  }#END for chr
}#END for indPair
```

